### 2018-03-05 chapter4/ms.tex

%UNLIKE IN A REGULAR TEX FILE, DON'T PUT ANY PREAMBLE MATERIAL HERE

\section*{Abstract}
Paper abstract goes here
\section{Paper Section}
Section stuff

\subsection{Paper Subsection}
\section{Paper Section}
Section stuff

\subsection{Paper Subsection}
\section{Paper Section}
Section stuff

\subsection{Paper Subsection}

\subsection{Paper Subsection}
\section{Final Paper Section}
Wrap up your paper here

\section*{Acknowledgments}
Put the acknowledgements from your paper here

%NOTE THE LACK OF A BIBLIOGRAPHY CALL IN THIS FILE. BIBLIOGRAPHY WORK HAPPENS OUTSIDE THE CHAPTER TEX FILES.