### 2018-03-08 abstract.tex
Outline abstract/Intro. Add rough title

\abstract{
Importance of joint GW-EM Astronomy. Development of contamination studies

GW Discoveries. GW170817. Our work on the optical modeling.

Lastly, LSST.
}