### 2018-03-13 chapter6/ms.tex
Start Adding Ch6

%UNLIKE IN A REGULAR TEX FILE, DON'T PUT ANY PREAMBLE MATERIAL HERE
\clearpage
\section*{Abstract}

\section{Introduction}
\label{sec:ch6_intro}
GW170817 new era stuff

Future searches, development of LIGO and additional detectors

LSST is well matched to LIGO follow-up

In this paper we do stuff and stuff

AB mags and Planck cosmology.

\section{Kilonova Models}
\label{sec:ch6_models}
Brief description of model construction + cite previous papers.

Describe actual models produced

Describe construction of red + blue models.

Describe Usage of GW170817

\section{Simulated Observations}
\label{sec:ch6_obs}

We consider two observing scenarios, survey and too

\subsection{OpsSim}
\label{sec:ch6_opssim}
Describe opssim

Our Setup

\subsection{SNANA/ToO}
\label{sec:ch6_snana}
Describe SNANA and our set up

Outline ToO strategies and considerations

\section{Analysis}
\label{sec:ch6_analysis}
Describe analysis goals. Detection and characterization of KN.

\subsection{OpsSim Results}
\label{sec:ch6_opssim_results}
We have such and such LC and find this rate

Some number are detected twice

Try to fit them?

\subsection{ToO Results}
\label{sec:ch6_too_results}
From SNANA we have such and such results.

Try to fit LC returned by SNANA

Look at things like ability to measure ejecta properties as a function of time investment

\section{Discussion and Conclusion}
\label{sec:ch6_conc}
We have done some LSST stuff

Main survey summary

ToO summary

LSST must be involved in direct GW follow-up, well matched and won't find many in main survey in a usable sense.

%NO BIB INFO