### 2018-05-03 conclusion/conclusion.tex
Add Conclusion

% No Preamble Material
Something about how new and great MMA is.

Blanket paragraph about GW170817 and how well-behaved it is.

From the optical and NIR data presented in this thesis alone, we can infer the following:
{\small \begin{enumerate}[leftmargin = 2.5cm]
\item[Strategy 1:] 
\end{enumerate}}

While this alone represents a phenomenal success for the first joint GW-EM detection, numerous open questions remain:

{\small \begin{enumerate}[leftmargin = 2.5cm]
\item[Strategy 1:] 
\end{enumerate}}

Hopeful outro paragraph about how neat all this will be in the future (LIGO increased sensitivities, New Instruments and discoveries await). Wau what a time to be alive.

% https://youtu.be/5sYtRpnOJzU?t=2401
% https://arxiv.org/pdf/1710.05931.pdf
% https://ui.adsabs.harvard.edu/?#abs/2017arXiv171005931M/references